---
title: "R Notebook"
output: html_notebook
---

### Create the dataframe
##Note: create multiple entries for each time category, so can do operations like min/max by group, and get actual results for the stat="bin" geom
##Note: the "levels" orders the categories for output (if you don't want them alpha order for ex)
```{r}
dat <- data.frame(
  time = factor(c("First Breakfast","Second Breakfast","Elevenses","Luncheon","Afternoon Tea","Dinner","Supper"), levels=c("First Breakfast","Second Breakfast","Elevenses","Luncheon","Afternoon Tea","Dinner","Supper")),
  total_bill = c(3,18.4,14.89, 17.23,5.02,9,8),
  tip = c(1,4,0,0,2,0,0),
  location = c("The Shire","Town","Friend's","Friend's","Town","The Shire","The Shire")
)
```

### Explore data

```{r}
# view dataframe contents
dat
# view first 4 rows in dataframe
head(dat,4)
# view last 11 rows in dataframe
# if you choose a # higher than row count, all rows will be displayed
tail(dat,11)
#view max of column
min(dat$total_bill)
#view min of column
min(dat$total_bill)
#view avg of column
mean(dat$total_bill)

#TIP: GET MIN/MAX OF ALL COLUMNS AT ONCE (INCLUDING STRINGS)
#Reference: https://stackoverflow.com/questions/13762364/how-to-find-the-minimum-value-of-a-column-in-r
apply(dat,2,min)
apply(dat,2,max)
#apply(dat,2,mean) #NOTE: can't do mean on a string
#BUT can you (?) get the avg of mult columns at once? do a for loop through a list of variable names?
#mean(dat$total_bill, dat$tip) # doesn't work
#https://stackoverflow.com/questions/30941504/calculate-mean-for-multiple-columns-in-data-frame
#this auto-detects which columns are numeric, then gets the value for each

print("Min")
sapply(Filter(is.numeric, dat), min)
print("Average")
sapply(Filter(is.numeric, dat), mean)
print("Max")
sapply(Filter(is.numeric, dat), max)
##there's a problem with sapply? doesn't handle continuous columns??? 
#source: https://stackoverflow.com/questions/5863097/selecting-only-numeric-columns-from-a-data-frame

##Get Unique Values per column
#FLAG: this is very slow for large datasets
unique(dat$time)
unique(dat$location)
##Next Step: how to get unique values of every column that is non-numeric in 1 step?
##Next Step: how to do a faster 
```

Next Step: Connect to Vertica and try doing the sapply min/max on real data
verify that there are no data type issues

```{r}
options(scipen=999)
dscc_scores_nv <- dbGetQuery(vertica, 
           'select personid, state_code
              ,support_score
              ,turnout_score
              ,gotv_score_bl
              ,earlyvote_score
              ,votebymail_score
           from dscc_scores_2018.dscc_all_scores_2018 
           where state_code = \'NV\';')

head(dscc_scores_nv)

#this will get min/max/avg for all numeric columns in a dataset
# na.rm = TRUE removes nulls from the results
print("Min")
sapply(Filter(is.numeric, dscc_scores_nv), min, na.rm = TRUE)
print("Average")
sapply(Filter(is.numeric, dscc_scores_nv), mean, na.rm = TRUE)
print("Max")
sapply(Filter(is.numeric, dscc_scores_nv), max, na.rm = TRUE)

#get counts of null values
sum(is.na(dscc_scores_nv$support_score))
```

#figure this out for the columns throwing NA above
```{r}
#create true/false vector of numeric columns
nums <- unlist(lapply(dscc_scores_nv, is.numeric))  
nums[1]
#create new df that consist of only numeric columns
dscc_scores_nv_num <- dscc_scores_nv[ , nums]
dscc_scores_nv_num

#get summary data off of new df
lapply(dscc_scores_nv_num, max)

```


Next step: Graphical Analysis QC
ex: score distribution

### Histograms

#next steps: fiddle with binwidth once i have real support score
Reference: http://www.sthda.com/english/wiki/ggplot2-histogram-plot-quick-start-guide-r-software-and-data-visualization
```{r}
library(ggplot2)

# Basic histogram - Hobbit
ggplot(dat, aes(x=total_bill)) + geom_histogram(binwidth=2) + theme_ss
# Tetris histogram - hobbit
ggplot(dat, aes(x=tip)) + geom_histogram(binwidth=2) + theme_ss

# Histogram - support score ---gradient fill needs `..x..` in the aes fill to work properly, I don't know why
ggplot(dat_csv, aes(x=support_score, fill= ..x.. )) + geom_histogram(binwidth=.005) + theme_bluelabs +
  #set colors to emerald scheme
  # scale_fill_gradient(low = colors_lowhigh[1], high=colors_lowhigh[2])
  #set only blue gradient, BL colors
  # scale_fill_gradient(low = bluelabs_colors[3], high=bluelabs_colors[5])
  #set to red/blue scale, BL colors
  scale_fill_gradient2(low=bluelabs_colors[7], mid="white", high=bluelabs_colors[5], midpoint=0.5)+
  # scale_fill_gradient(low='blue', high='yellow')
  labs(title="Support Score Distribution",
       y = element_blank(),
       x = element_blank())+
  #remove legend
  guides(fill=FALSE)+
  #remove y-axis labels
  theme(axis.text.y = element_blank())

# Histogram - ev score
ggplot(dat_csv2, aes(x=ev_score)) + geom_histogram(binwidth=.05) + theme_ss
```

### Density Plots --try this with different, real data
```{r}
# Histogram - support score ---gradient fill needs `..x..` in the aes fill to work properly, I don't know why
ggplot(dat_csv, aes(x=support_score)) + geom_density() + theme_bluelabs +
 labs(title="Support Score Distribution\nOR\nA Poorly Baked Loaf of Bread\n?",
       y = element_blank(),
       x = element_blank())+
  guides(fill=FALSE)+
  theme(axis.text.y = element_blank())

#needed for dollar_format below
library(scales)

#density plot: hobbit
ggplot(dat, aes(x=total_bill)) +geom_density() + theme_ss
ggplot(dat, aes(x=tip)) +geom_density() + theme_ss +
  annotate(geom = "text", x=.5, y=(.2), label='Takeaway:\nHobbits Are\nBad Tippers', size = 4, fontface=1.2, color=colors_ss[4])+
   labs(title="Tipping Habits of Hobbits",
       y = element_blank(),
       x = element_blank())+
  theme(axis.text.y = element_blank())+
  theme(axis.text.x = element_text(size=12))+
  scale_x_continuous(labels=dollar_format(prefix="€"))
```

# GEt structure of data frame
```{r}
str(dat_csv3)
summary(dat_csv3)
```

